### 2014-10-18 TrialFile.Rmd
Tried making a table for the proposal document

---
title: "Untitled"
author: "Sam"
date: "October 18, 2014"
output: html_document
---

This is an extra file for trying out features including making maps for the proposal.

```{r echo=FALSE, cache=TRUE}
install.packages('maptools')
library(maptools)
install.packages('maps')
library(maps)
install.packages('mapproj')
library(mapproj)
install.packages('ggmap')
library(ggmap)
install.packages('mapdata')
library(mapdata)

library(sp)
library(lattice)
install.packages('latticeExtra')
library(latticeExtra)
library(colorspace)

# Create maps of Sweden, France, and the UK using maps and mapdata packages
map('worldHires', 'sweden', col="gray90", fill=TRUE)

map('worldHires',
    c('UK', 'Ireland', 'Isle of Man','Isle of Wight','Wales:Anglesey'),
    xlim=c(-11,3), ylim=c(49,60.9),
    col="gray90", fill=TRUE)

map('worldHires','France', xlim=c(-8.2,8.5), ylim=c(42.2,51.1),
    col="gray90", fill=TRUE)
```

<table>
<tr class="header">
<th align="left"> </th>
<th align="left"> </th>
</tr>
<tr class="odd">
<td align="left">Finance</td>
<td align="left">Government Policy</td>
<td align="left">Entrepreneurial Education and Training</td>
<td align="left">Research and Development Transfer (R & D)</td>

</tr>
<tr class="even">
<td align="left">Commercial and professional infrastructure</td>
<td align="left">Internal Market openness</td>
<td align="left">Physical infrastructure and services</td>
<td align="left">Cultural and social norms </td>
</tr>
</table>