### 2014-10-28 WebScraping1.Rmd
Code for webscraping GEM

---
title: "WebScraping1"
author: "Sam"
date: "October 27, 2014"
output: html_document
---

```{r}

library(httr)
library(dplyr)
install.packages("XML")
library(XML)

URL <- 'http://www.bbc.com/sport/winter-olympics/2014/medals/countries'

# Get and parse all tables on page
tables <- URL %>% GET() %>%
  content(as = 'parsed') %>%
  readHTMLTable()

# Gather content and parse all tables (only one on this page)
names(tables)
medals <- as.data.frame(tables)

### Clean ###
# Drop unwanted variables
medals

MedalsTable <- tables[]

head(MedalsTable)
summary(MedalsTable)
arrange <- arrange()
```

```{r}

# Install Foreign for accessing STATA file
install.packages("foreign")
library(foreign)
library(help=foreign)

# Load data locally
dataSPSS<-read.spss("/Users/samdundon/Desktop/2010GEMNES.sav", to.data.frame=TRUE)

### Clean ###

head(dataSPSS)
head(dataSPSS)[1,1:387]

```