### 2014-11-14 GEMDataSource.Rmd
Edits to markdown for consistency, organized repository, and created new README

---
title: "Education and the Rate of Entrepreneurship - Data Gathering and Analysis"
author: "Samuel Dundon and Thierry Racaud"
date: "November 14, 2014"
output: html_document
---
***

*Introduction to Collaborative Social Science Data Analysis*

#### Research Question

To what extent do the behavioral characteristics of education systems - individualistic, group-oriented, self-esteem reinforcing - affect the rate of entrepreneurship in OECD countries?  
<br>

#### Causal Chain

Education system / Teaching reinforcing self-esteem ---> higher degree of individual self-esteem ---> higher probability an individual will become an entrepreneur ---> higher rate of entrepreneurship at a country-level.
<br>

##### Dependent Variables: Explaining the Country-level Rate of Entrepreneurship

We have examined three dimensions of entrepreneurship to prepare for this stage of the project, established entrepreneurship in the country, nascent entrepreneurship (less than three months), and entrepreneurial intention. We will study each of these successively as we may well observe different effects.

To measure these three dimensions of entrepreneurship, we have used three [Global Entrepreneurship Monitor (GEM)](http://www.gemconsortium.org) indicators:

- Established Business Ownership Rate (Percentage) 
- Nascent Entrepreneurship Rate (Percentage)
- Entrepreneurial intention Rate (Percentage)

##### Key Explanatory Variable: Degree of Self-esteem Reinforcement within the Education System

As Inglehart has shown how personal values are shaped before the age of 20, we will try to show how compulsory education (from 6 to 16 years of age) shapes individuals, especially in terms of self-esteem. To do so, we decided to incorporate GEM indicators that relate to self-esteem and confidence. First, we included GEM survey data on the average "perceived capabilities" of an individual (18-64) in a given country to start a business. Second, we included additional GEM data on how the "fear of failure" could deter an average individual (18-64) in a country to become an entrepreneur. Both of these data sets indicate the average percentage for each characteristic at the country-level.

Going forward, we are considering adding one of the six questions used by GEM to measure the importance of entrepreneurial education at country-level in its NES Survey. Item D01 states: "In my country, teaching in primary and secondary education encourages creativity, self-sufficiency, and personal initiative."

##### Control Variables: Controlling for Classic Determinants of Entrepreneurship

To control for the general state of the economy, we included annual GDP Growth from the World Bank [World Development Indicators](http://data.worldbank.org/data-catalog/world-development-indicators) as a control variable in our regression equations.  In addition, we included a GEM indicator called "Perceived Opportunities." For each country, it is the average percentage of individuals who believe there are opportunities to start a business in their area. Since we look at average individual self-esteem in every OECD country, we believe it is appropriate to look at the opportunities perceived by the average 18-64-year-old individual in these countries.

In the future, we may revise this to include lagged GDP growth in recognition of the time taken for governments and individuals to react to the state of the economy, and are planning to include control variables from GEM. Specifically, we intend to control for the classic determinants of entrepreneurship -- economic and structural determinants. To do so, we will add the following nine key entrepreneurship conditions as defined in the GEM [National Experts Survey (NES)](http://http://www.gemconsortium.org/Data-Collection).

1. Finance  
2. Government policies  
3. Government programs  
4. Entrepreneurial Education and Training  
5. Research and Development Transfer (R & D)  
6. Commercial and professional infrastructure  
7. Internal Market openness  
8. Physical infrastructure and services  
9. Cultural and social norms

The following equations were used to determine the relationship between our dependent and explanatory variables:

<br>

1. $PercentIntention = PercentPerceived + PercentFear*PercentOpportunities + AnnualGDPGrowth$

2. $PercentNascent = PercentPerceived + PercentFear*PercentOpportunities + AnnualGDPGrowth$

3. $PercentOwner = PercentPerceived + PercentFear*PercentOpportunities + AnnualGDPGrowth$  
<br>

The variables *Percent Fear* and *Percent Opportunities* were included as an interaction term $PercentFear*PercentOpportunities$ because of the design of the survey. PercentFear is:

>The Percentage of 18-64 population with positive perceived opportunities who indicate that fear of failure would prevent them from setting up a business.

<br>

#### Data Collection and Cleaning

To collect the data necessary to run these regression equations, we downloaded *.csv* data from GEM and used the WDI API to obtain the necessary annual GDP growth of OECD Countries.

**Necessary R Packages**

The following packages were needed to complete the data collection portion of this assignment:

* repmis  
* tidyr  
* reshape  
* plyr  
* WDI

**GEM Data**

GEM data was embedded in an interactive database and was difficult to extract automatically without writing a program. To simplify this process, we downloaded the necessary *.csv* files and saved them in our GitHub repository (to reduce complexity, we deleted the first three unnecessary rows before saving).  We then linked to the RAW version of the files and added the code required to load them from *R*.  The repmis package is required to load these documents from the repository.

Cleaning the GEM data required deleting unnecessary characters in the year columns, changing the years from columns to rows (changing the data from wide to long), and changing the column names to describe which data was being analyzed.  In the process of doing this we created new data sets for each of these survey questions, which were called C*dataset* (clean followed by the data set name). This was done for each of the six GEM data sets before merging these sets two at a time to create the data set *FinalGEM*.

For all 33 OECD countries we should have 13 observations, one per year from 2001 to 2013, for each variable. However, there are a few missing values, especially regarding the perceived opportunities for some countries, namely Estonia, Luxembourg and Slovak Republic.

**World Bank Data**

As mentioned above, one of our control variables, annual GDP growth, was obtained using the WDI API from the World Bank.  Using this API, we accessed GDP data from all the OECD countries directly from *R*. To clean this data, we used *plyr* to change the variable names then recoded the U.S. and South Korea country names to maintain consistency with the GEM data set and dropped an unnecessary column that was not included with the GEM data.

Finally, we merged the annual GDP growth data with the *FinalGEM* data set to create *CompleteDataset*.  This is the data set used to run the regressions and generate the charts and figures outlined in this document.  To reduce complexity in the environment we also included a command to remove unnecessary objects in the environment.

<br>
```{r echo =FALSE, cache=TRUE, include=FALSE, results='asis'}

# Set working directory
setwd('~/Assign2Education/')

# Library required packages repmis, ridy, reshape2, plyr, WDI, and stargazer packages
library(repmis)
library(tidyr)
library(reshape2)
library(plyr)
library(WDI)
library(stargazer)

# Load citation package
pkgs <- c('repmis', 'tidyr', 'reshape2', 'plyr', 'WDI')
repmis::LoadandCite(pkgs, file= 'DataRpackageCitations.bib')

#### Import data from GEM Survey ####

# Import GEM fear of failure data
FearFail <- repmis::source_data("https://raw.githubusercontent.com/SamDund/Assign2Education/master/GEMData/GEM_FearFail.csv", header = TRUE, cache = FALSE)

# Import GEM estblished entrepreneurship - current owner data
CurrentOwner <- repmis::source_data("https://raw.githubusercontent.com/SamDund/Assign2Education/master/GEMData/GEM_CurrentOwner.csv", header = TRUE, cache = FALSE)

# Import GEM entrepreneurial intention data
EntreIntention <- repmis::source_data("https://raw.githubusercontent.com/SamDund/Assign2Education/master/GEMData/GEM_EntreIntention.csv", header = TRUE, cache = FALSE)

# Import GEM nascent entrepreneur data
NascentEntre <- repmis::source_data("https://raw.githubusercontent.com/SamDund/Assign2Education/master/GEMData/GEM_NascentEntre.csv", header = TRUE, cache = FALSE)

# Import GEM perceived capabilities data
PerceivedCapabilities <- repmis::source_data("https://raw.githubusercontent.com/SamDund/Assign2Education/master/GEMData/GEM_PerceivedCapabilities.csv", header = TRUE, cache = FALSE)

# Import GEM perceived opportunities data as a proxy for potential control variables
PerceivedOpportunities <- repmis::source_data("https://raw.githubusercontent.com/SamDund/Assign2Education/master/GEMData/PerceivedOpportunities.csv", header = TRUE, cache = FALSE)

#### Clean data using tidy ####

### Clean FearFail ###
# Percentage of 18-64 population with positive perceived opportunities who indicate that fear of failure would prevent them from setting up a business.

# Remove X's in FearFail year columns
names(FearFail) <- gsub("X", "", names(FearFail))

# Change years from columns to rows and create CFearFail
CFearFail <- melt(data = FearFail,
                  id.vars = c("Country"),
                  measure.vars = c("2001", "2002", "2003", "2004", "2005", "2006",
                                   "2007", "2008", "2009", "2010", "2011", "2012",
                                   "2013"))

# Rename columns
CFearFail <- plyr::rename(x = CFearFail, 
                                      replace = c("variable" = "Year",
                                                  "value" = "PercentFear"))

### Clean CurrentOwner ###
# Percentage of 18-64 population who are currently owner-manager of an established business, i.e., owning and managing a running business that has paid sallaries, wages, or any other payments to the owners for more than 42 months

# Remove X's in CurrentOwner year columns
names(CurrentOwner) <- gsub("X", "", names(CurrentOwner))

# Change years from columns to rows and create CCurrentOwner
CCurrentOwner <- melt(data = CurrentOwner,
                  id.vars = c("Country"),
                  measure.vars = c("2001", "2002", "2003", "2004", "2005", "2006",
                                   "2007", "2008", "2009", "2010", "2011", "2012",
                                   "2013"))

# Rename columns
CCurrentOwner <- plyr::rename(x = CCurrentOwner, 
                                      replace = c("variable" = "Year",
                                                  "value" = "PercentOwner"))

### Clean EntreIntention ###
# Percentage of 18-64 population (individuals involved in any stage of entrepenerurial activity excluded) who intend to start a business within three years.

# Remove X's in EntreIntention year columns
names(EntreIntention) <- gsub("X", "", names(EntreIntention))

# Change years from columns to rows and create CEntreIntention (No 2001 data available)
CEntreIntention <- melt(data = EntreIntention,
                  id.vars = c("Country"),
                  measure.vars = c("2002", "2003", "2004", "2005", "2006",
                                   "2007", "2008", "2009", "2010", "2011", "2012",
                                   "2013"))

# Rename columns
CEntreIntention <- plyr::rename(x = CEntreIntention, 
                                      replace = c("variable" = "Year",
                                                  "value" = "PercentIntention"))

### Clean NascentEntre ###
# Percentage of 18-64 population who are currently a nascent entrepreneur, i.e., actively involved in setting up a business they will own or co-own; this business has not paid salaries, wages, or any other payments to the owners for more than three months.

# Remove X's in NascentEntre year columns
names(NascentEntre) <- gsub("X", "", names(NascentEntre))

# Change years from columns to rows and create CNascentEntre
CNascentEntre <- melt(data = NascentEntre,
                  id.vars = c("Country"),
                  measure.vars = c("2001", "2002", "2003", "2004", "2005", "2006",
                                   "2007", "2008", "2009", "2010", "2011", "2012",
                                   "2013"))

# Rename columns
CNascentEntre <- plyr::rename(x = CNascentEntre, 
                                      replace = c("variable" = "Year",
                                                  "value" = "PercentNascent"))

### Clean PerceivedCapabilities ###
# Percentage of 18-64 population who believe to have the required skills and knowledge to start a business

# Remove X's in PerceivedCapabilities year columns
names(PerceivedCapabilities) <- gsub("X", "", names(PerceivedCapabilities))

# Change years from columns to rows and create CPreceivedCapabilities
CPerceivedCapabilities<- melt(data = PerceivedCapabilities,
                  id.vars = c("Country"),
                  measure.vars = c("2001", "2002", "2003", "2004", "2005", "2006",
                                   "2007", "2008", "2009", "2010", "2011", "2012",
                                   "2013"))

# Rename columns
CPerceivedCapabilities <- plyr::rename(x = CPerceivedCapabilities, 
                                      replace = c("variable" = "Year",
                                                  "value" = "PercentPerceived"))

### Clean PerceivedOpportunities ###
# Percentage of 18-64 population who see good opportunities to start a firm in the area where they live

# Remove X's in PerceivedOpportunities year columns
names(PerceivedOpportunities) <- gsub("X", "", names(PerceivedOpportunities))

# Change years from columns to rows and create CPreceivedOpportunities
CPerceivedOpportunities<- melt(data = PerceivedOpportunities,
                  id.vars = c("Country"),
                  measure.vars = c("2001", "2002", "2003", "2004", "2005", "2006",
                                   "2007", "2008", "2009", "2010", "2011", "2012",
                                   "2013"))

# Rename columns
CPerceivedOpportunities <- plyr::rename(x = CPerceivedOpportunities, 
                                      replace = c("variable" = "Year",
                                                  "value" = "PercentOpportunities"))

#### Merge the data ####
MergedGEM <- merge(CCurrentOwner, CEntreIntention,
                   union("Country", "Year"),
                   all = TRUE)

MergedGem2 <- merge(CFearFail, CNascentEntre,
                   union("Country", "Year"),
                   all = TRUE)

MergedGem3 <- merge(CPerceivedCapabilities, MergedGEM,
                   union("Country", "Year"),
                   all = TRUE)

MergedGem4 <- merge(MergedGem2, MergedGem3,
                   union("Country", "Year"),
                   all = TRUE)

FinalGEM <- merge(MergedGem4, CPerceivedOpportunities,
                   union("Country", "Year"),
                   all = TRUE)

# Convert percent columns from character to numeric
FinalGEM[,"PercentPerceived"] <- as.numeric(FinalGEM[,"PercentPerceived"])
FinalGEM[,"PercentFear"] <- as.numeric(FinalGEM[,"PercentFear"])
FinalGEM[,"PercentNascent"] <- as.numeric(FinalGEM[,"PercentNascent"])
FinalGEM[,"PercentOwner"] <- as.numeric(FinalGEM[,"PercentOwner"])
FinalGEM[,"PercentIntention"] <- as.numeric(FinalGEM[,"PercentIntention"])
FinalGEM[,"PercentOpportunities"] <- as.numeric(FinalGEM[,"PercentOpportunities"])

#### Import WDI Annual GDP Growth Data ####
GDPGrowth = WDI(indicator='NY.GDP.MKTP.KD.ZG', country=c('MX','CA','US','FR','AU','AT','BE',
                                                         'CL','CZ','DK','EE','FI','DE','HU', 
                                                         'IS','IE','IL','IT','JP','KR','LU',
                                                         'NL','NZ','NO','PL','PT','SK','SI',
                                                         'ES','SE','CH','TR','GB','US','GR'), 
                start=2001, end=2013)
# Rename columns
GDPGrowth <- plyr::rename(x = GDPGrowth, 
                                      replace = c("year" = "Year",
                                                  "NY.GDP.MKTP.KD.ZG" = "AnnualGDPGrowth",
                                                  "country" = "Country"))

# Recode country == "United States" to "United States of America" and Korea, Rep. to Korea (South)
GDPGrowth$Country[GDPGrowth$Country == "United States"] <- "United States of America"

# Recode country == "Korea, Rep."" to "Korea (South)"
GDPGrowth$Country[GDPGrowth$Country == "Korea, Rep."] <- "Korea (South)"

# Remove iso2c data to ease merge with FinalMerge data
GDPGrowth$iso2c <- NULL

# Merge GDPGrowth and FinalGEM
CompleteDataset <- merge(FinalGEM, GDPGrowth,
                   union("Country", "Year"),
                   all = TRUE)

# Remove unncessary data from environment
remove(CCurrentOwner, CEntreIntention, CFearFail, CNascentEntre, CPerceivedCapabilities, CPerceivedOpportunities, CurrentOwner, EntreIntention, FearFail, MergedGEM, MergedGem2, MergedGem3, MergedGem4, PerceivedCapabilities, PerceivedOpportunities, NascentEntre)

```

#### Data Analysis

To review, the three regression equations run for this assignment are the following:

1. $PercentIntention = PercentPerceived + PercentFear*PercentOpportunities + AnnualGDPGrowth$

2. $PercentNascent = PercentPerceived + PercentFear*PercentOpportunities + AnnualGDPGrowth$

3. $PercentOwner = PercentPerceived + PercentFear*PercentOpportunities + AnnualGDPGrowth$

To examine the relationship between these variables, we used least square estimation, the results of which can be found in the following table.  The package *stargazer* was required to generate this table.  As evidenced by the equation results, the individual variables selected are statistically significant, but the model only explains a relatively small degree of the variation in the dependent variable.  For example, the adjusted $R^2$ of regression equation one is 0.29, the adjusted $R^2$ of regression equation two is 0.38, and the adjusted $R^2$ is 0.13.  
<br>

```{r echo =FALSE, cache=TRUE, include=FALSE, results='asis'}

# Run linear regression R1
R1 <- lm(PercentIntention ~ PercentPerceived + PercentFear*PercentOpportunities + AnnualGDPGrowth, data = CompleteDataset)

# Obtain summary statistics of R1
summary(R1)

# Run linear regression R2
R2 <- lm(PercentNascent ~ PercentPerceived + PercentFear*PercentOpportunities + AnnualGDPGrowth, data = CompleteDataset)

# Obtain summary statistics of R2
summary(R2)

# Run linear regression R3
R3 <- lm(PercentOwner ~ PercentPerceived + PercentFear*PercentOpportunities + AnnualGDPGrowth, data = CompleteDataset)

# Obtain summary statistics of R3
summary(R3)
```

```{r echo = FALSE, cache=TRUE, include=TRUE, results='asis'}

# Create regression output table
labels <- c('Percent Perceived', 'Percent Fear',
            'Annual GDP Growth', 'Percent Opportunities', 'Percent Fear x Percent Opportunities','(Intercept)')

stargazer::stargazer(R1, R2, R3, covariate.labels = labels, 
                     title = 'Linear Regression Estimates', digits = 2, type = 'html')
```
<br>

**Simulation Package and Cleaning**

In addition to the output table, a simulation is helpful to visually illustrate the relationship the probability of obtaining a given value in the dependent variables conditional on change in the independent variable.  We used the *Zelig* package to run a least square regression and generate a simulation of equation 1.  Prior to running this regression, however, we ran the na.omit command to remove omitted variables from the equation.

To examine the probabilities of achieving given outcomes, we generated three simulations for the first regression equation.  To demonstrate this method, we have included one simulation below.

**Process**

First, we estimated a least square equation with variation in the variable *PercentPerceived* of 20-100 percent. The ranges for variance in the independent variables was chosen by identifying the best range for capturing the largest degree of variance for the model.  The first simulation is below.  The y axis shows variation in *PercentIntention* conditional on change in *PercentPerceived* and the x axis shows variation in *PercentPerceived*. The results may be interpreted as showing that people who believe they have the knowledge required to start a business may or may not do so (with a high degree of variation), while people who do not believe they have this knowledge are consistently unlikely to start a business (there is little variation).

```{r echo = FALSE, cache=TRUE, include=FALSE, results='asis'}

# Library Zelig (can be buggy if loaded to early)
library(Zelig)

# Omit omitted data from simulation
CompleteData_omit <- na.omit(CompleteDataset)

# Run a least squares regression with Zelig
Z1 <- zelig(PercentIntention ~ PercentPerceived + PercentFear*PercentOpportunities + AnnualGDPGrowth, data = CompleteData_omit, model = "ls")
```

```{r echo = FALSE, cache=TRUE, include=TRUE, results='asis' }

# Set range of PercentPerceived values in simulation
x.out2 <- setx(Z1, PercentPerceived = 20:100)

# Run simulation (default 1000 trials)
simz2 <- sim(Z1, x = x.out2)

# Adjust margins
par(mar=c(2,2,2,2))

# Plot simulation
plot(simz2)

```

<br>

#### Summary

The impact of cultural factors in education on the rate of entrepreneurship in a given country is an important and under-researched topic that should be considered by policymakers constructing public education budgets. Through this project, we hope to encourage the consideration of these cultural factors by providing a qualitative and quantitative method for understanding the impact of these factors. Although several challenges remain to compiling the indexes discussed in this report, we are confident we can produce a helpful framework for examining and understanding these factors.