### 2014-11-29 FromEducationToEntrepreneurship.Rmd
Clean Rmd

---
title: "From Education To Entrepreneurship"
author: "Thierry Racaud"
date: "29 novembre 2014"
output: html_document
---

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this: