### 2014-11-29 FromEducationToEntrepreneurship.Rmd
Cleaning

---
title: "From Education To Entrepreneurship"
author: "Sam Dundon and Thierry Racaud"
date: "29 novembre 2014"
output: html_document
---

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

* * * 
```{r echo=FALSE, cache=TRUE, include=FALSE}

# Set working directory
setwd('~/Assign2Education/Proposal')

# Install necessary packages

# Install rempis from Christpher Gandrud's website
devtools::install_github('christophergandrud/repmis')
library(repmis)

# Load citation package
pkgs <- c('repmis')
repmis::LoadandCite(pkgs, file= 'RpackageCitations.bib')

```