### 2014-12-01 Source/Untitled.Rmd

---
title: "Untitled"
author: "Sam"
date: "November 29, 2014"
output: html_document
---

```{r}

library(RCurl)
library(devtools)
install_github(c('rstudio/htmltools',
            'ramnathv/htmlwidgets',
            'christophergandrud/networkD3'))
library(htmltools)
library(htmlwidgets)
library(networkD3)

# Plot
forceNetwork(Links = MisLinks, Nodes = MisNodes,
            Source = "source", Target = "target",
            Value = "value", NodeID = "name",
            Group = "group", opacity = 0.8)
```

```{r}
library(htmltools)
library(htmlwidgets)
library(networkD3)
library(RCurl)
library(devtools)
library(repmis)

Data1 <- repmis::source_data("https://raw.githubusercontent.com/SamDund/Assign2Education/master/GEMData/CompleteData.csv", header = TRUE, cache = FALSE)

DLinks <- Data1toDF(Data1$PercentPerceived = Energy,
                    array = "links")
DNodes <- Data1toDF(Data1$Country = Energy,
                    array = "nodes")

```