### 2014-12-02 Presentation/PresMisc/MiscPresStuff.Rmd
Reorganized repository and created histogram

---
title: "Untitled"
author: "Sam"
date: "December 2, 2014"
output: ioslides_presentation
---

## R Markdown
```{r}
# Create googleVis chart
line_plot <- gvisLineChart(D1)
plot(line_plot)

# Create vector of OECD countries
places <- c('Australia', 'Austria', 'Belgium', 'Canada', 'Chile', 'Czech Republic', 'Denmark', 'Estonia', 'Finland', 'France', 'Germany', 'Greece', 'Hungary', 'Iceland', 'Ireland', 'Israel', 'Italy', 'Japan', 'Korea', 'Luxemburg', 'Mexico', 'Netherlands', 'New Zealand', 'Norway', 'Poland', 'Portugal', 'Slovak Republic', 'Slovenia', 'Spain', 'Sweden', 'Switzerland', 'Turkey', 'United Kingdom', 'United States')

# Find latitude and longitude for OECD countries
geocode(places)

# Generate a map
qmap(location = 'Australia', zoom = 4)
```
## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Code and Output

```{r}
summary(cars)
```

## Slide with Plot

```{r, echo=FALSE}
plot(cars)
```