### 2014-12-03 Presentation/PresMisc/MiscPresStuff.Rmd

---
title: "Untitled"
author: "Sam"
date: "December 2, 2014"
output: ioslides_presentation
---

## R Markdown
```{r}
# Create googleVis chart
line_plot <- gvisLineChart(D1)
plot(line_plot)

# Create vector of OECD countries
places <- c('Australia', 'Austria', 'Belgium', 'Canada', 'Chile', 'Czech Republic', 'Denmark', 'Estonia', 'Finland', 'France', 'Germany', 'Greece', 'Hungary', 'Iceland', 'Ireland', 'Israel', 'Italy', 'Japan', 'Korea', 'Luxemburg', 'Mexico', 'Netherlands', 'New Zealand', 'Norway', 'Poland', 'Portugal', 'Slovak Republic', 'Slovenia', 'Spain', 'Sweden', 'Switzerland', 'Turkey', 'United Kingdom', 'United States')

# Find latitude and longitude for OECD countries
geocode(places)

# Generate a map
qmap(location = 'Australia', zoom = 4)
```
## Slide with Bullets

```{r echo =FALSE, cache=FALSE, include=TRUE, results='asis', warning=FALSE, error=FALSE, message=FALSE}
library("ggvis")
# Interactive histgram (isn't interactive in knitted docs)
D4 %>% 
  ggvis(~mean, fill := "#2c7fb8") %>% 
  layer_histograms(width =  input_slider(0, 2, step = 0.10, label = "width"),
                   center = input_slider(0, 2, step = 0.05, label = "center"))

```

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Code and Output

```{r}
summary(cars)
```

## Slide with Plot

```{r, echo=FALSE}
# sankeyNetwork
library(RCurl)
library(devtools)
install_github(c('rstudio/htmltools',
            'ramnathv/htmlwidgets',
            'christophergandrud/networkD3'))
library(htmltools)
library(htmlwidgets)
library(networkD3)

URL <- "https://raw.githubusercontent.com/christophergandrud/d3Network/sankey/JSONdata/energy.json"
Energy <- getURL(URL, ssl.verifypeer = FALSE)

# Convert to data frame
EngLinks <- JSONtoDF(jsonStr = Energy, array = "links")
EngNodes <- JSONtoDF(jsonStr = Energy, array = "nodes")

```

```{r}
# Histogram
D4 %>% ggvis(~mean, fill := "#2c7fb8") %>% layer_histograms(width = 0.2)

# Library stargazer
library(stargazer)

# Create new DF for PerceivedOpportunities
D5 <- PerceivedOpportunities

# Eliminate X's in header
names(D5) <- gsub("X", "", names(D5))  
D5$new.col <- NULL

# Create descriptive statistics table
T1 <- stargazer(D5, type = 'html', column.sep.width = "20pt", digits = 2)

##################

###################################################

library(googleVis)
library(reshape2)

# Melt Opp1 (Perceived opportunities 2001:2013 OECD)
dfChart1 <- melt(data = Opp2,
                  id.vars = c("Country", "iso2c"),
                  measure.vars = c("2001", "2002", "2003", "2004", "2005", "2006",
                                   "2007", "2008", "2009", "2010"))

# Change column name of value and variable
names(dfChart1) <- gsub("value", "WayofTeaching", names(dfChart1))
names(dfChart1) <- gsub("variable", "Year", names(dfChart1))

library(rCharts)

# rChart
Try1 <- transform(dfChart1, Year = as.character('Year'))

sub_aust <- subset(Try1, Country == "Australia")

m1 <- mPlot(y = "WayofTeaching", x = "Year", type = "Line", data = Opp3)

m1$set(pointSize = 0, lineWidth = 1)

m1$print('chartTest')
```