### 2014-12-04 Presentation/EdEntrePres.Rmd
Updated slides

---
title: "The Impact of Education <br>
on Entrepreneurship"
author: "Sam Dundon and Thierry Racaud"
date: "December 2, 2014"
output: ioslides_presentation
---

## Research Interest

*Text from Thierry*

## Causal Chain

## Where our intuition came from

Variable | France | UK
-------------|---------------|---------
Fear of Failure (GEM 2012) | 43% | 36%
Perceived Capabilities (GEM 2012) | 36% | 47%
Sense of belonging (PISA 2012) | 47% | 79%

## What is PISA

- PISA: "Programme for International Student Assessment is a triennial international survey which aims to evaluate education systems worldwide by testing the skills and knowledge of 15 year old students."

- Data availability prevented us from using PISA data for this project

## What is GEM

- GEM: Global Entrepreneurship Monitor project is an annual assessment of the entrepreneurial activity, aspirations, and attitudes of individuals across a wide range of countries.

-

## Estimation Strategy

Variable | Indicator 
-------------|---------------
Y Dependent Vaiable | Rate of Entrepreneurship (Intention, Nascent, Established)
X1 Explanatory Variable | Way of Teaching
X2 Control Variable | Perceived Capabilities
X3 Control Variable | Fear of Failure / Perceived Opportunities
X4 Control Variable | Perceived Opportunities

## Explaining Way of Teaching

*How is it measured*

## Variation in how well Education encourages Self Confidence in OECD Countries

```{r echo =FALSE, cache=FALSE, include=TRUE, results='asis', warning=FALSE, error=FALSE, message=FALSE}
# Import GEM data frame
WayofTeaching <- repmis::source_data("https://raw.githubusercontent.com/SamDund/Assign2Education/master/GEMData/WayofTeaching", header = TRUE, cache = FALSE, na.strings = "NA")

# Drop unnecessary column 1
WayofTeaching$X <- NULL

# Change data from long to wide format
WT <- reshape(WayofTeaching, idvar = c("iso2c", "Country"), timevar = "Year", direction = "wide")

# Delete variable name in Year columns
names(WT) <- gsub("WayofTeaching.", "", names(WT))

# Duplicate df for future use
WT2 <- WT

# Generate a new column with mean of all other columns
WT$mean <- rowMeans(subset(WT, select = c("2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008",
                                          "2009", "2010")), na.rm = TRUE)

# Drop year columns, leaving only mean column
WT$'2001' <- NULL
WT$'2002' <- NULL
WT$'2003' <- NULL
WT$'2004' <- NULL
WT$'2005' <- NULL
WT$'2006' <- NULL
WT$'2007' <- NULL
WT$'2008' <- NULL
WT$'2009' <- NULL
WT$'2010' <- NULL
WT$'2011' <- NULL
WT$'2012' <- NULL
WT$'2013' <- NULL
WT$row.names <- NULL

# Create 
WT[, 3] <- round(log(WT[, 3]), digits = 2)
```

```{r echo = FALSE, cache= TRUE, include= TRUE, results= 'asis'}
library(googleVis)

# Plot
WT_map <- gvisGeoChart(WT, locationvar = 'iso2c',
                      colorvar = 'mean',
                      options = list(
                        colors = "['#ece7f2', '#2b8cbe']"
))

print(WT_map, tag = 'chart')

```

## Variation in Impact of Education on Self Confidence
```{r echo =FALSE, cache=TRUE, include=FALSE, results='asis', warning=FALSE, error=FALSE, message=FALSE}
library("ggplot2")
library("reshape2")

# Import df
Wdf <- repmis::source_data("https://raw.githubusercontent.com/SamDund/Assign2Education/master/GEMData/CompleteData.csv")

# Delete unnessesary columns
Wdf$X <- NULL
Wdf$PercentFear <- NULL
Wdf$PercentNascent <- NULL
Wdf$PercentPerceived <- NULL
Wdf$PercentOwner <- NULL
Wdf$PercentIntention <- NULL
Wdf$PercentOpportunities <- NULL

# Create country vector to subset
myvars2 <- c('US','FR','DE','JP','GB')

# Subset WT2
Wdf2 <- subset(Wdf, iso2c %in% myvars2)

```

```{r echo =FALSE, cache=TRUE, include=TRUE, results='asis', warning=FALSE, error=FALSE, message=FALSE}

# Create plot
ggplot(Wdf2,aes(x = Year, y = WayofTeaching, colour = Country, group = Country)) + geom_line()

```

## Perceived Capabilities

*What about it*

## Perceived Capabilities
```{r echo =FALSE, cache=TRUE, include=FALSE, results='asis'}

# Library required packages
library(googleVis)
library(countrycode)
library(repmis)

# Import GEM data frame
PerceivedCapabilities <- repmis::source_data("https://raw.githubusercontent.com/SamDund/Assign2Education/master/GEMData/GEM_PerceivedCapabilities.csv", header = TRUE, cache = FALSE, na.strings = "-")

## Add country codes and place new column at beginning of the data frame
# Create a vector of country codes
C2 <- countrycode(PerceivedCapabilities$Country, "country.name", "iso2c")

# Create a new column and insert the country code vector
PerceivedCapabilities$new.col <- C2

# Reorder columns so country code is first
C3 <- PerceivedCapabilities[,c(15,1:14)]

# Rename first column
C4 <- plyr::rename(x = C3, 
                   replace = c("new.col" = "iso2c"))

# Remove X's in D4 year columns
names(C4) <- gsub("X", "", names(C4))

# Set columns as numeric
C4[,"2001"] <- as.numeric(C4[,"2001"])
C4[,"2002"] <- as.numeric(C4[,"2002"])
C4[,"2003"] <- as.numeric(C4[,"2003"])
C4[,"2004"] <- as.numeric(C4[,"2004"])
C4[,"2005"] <- as.numeric(C4[,"2005"])
C4[,"2006"] <- as.numeric(C4[,"2006"])
C4[,"2007"] <- as.numeric(C4[,"2007"])
C4[,"2008"] <- as.numeric(C4[,"2008"])
C4[,"2009"] <- as.numeric(C4[,"2009"])
C4[,"2010"] <- as.numeric(C4[,"2010"])
C4[,"2011"] <- as.numeric(C4[,"2011"])
C4[,"2012"] <- as.numeric(C4[,"2012"])
C4[,"2013"] <- as.numeric(C4[,"2013"])

# Save cleaned df
Cap1 <- C4

# Generate a new column with mean of all other columns
C4$mean <- rowMeans(subset(C4, select = c("2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008",
                                          "2009", "2010", "2011", "2012", "2013")), na.rm = TRUE)

# Drop year columns, leaving only mean column
C4$'2001' <- NULL
C4$'2002' <- NULL
C4$'2003' <- NULL
C4$'2004' <- NULL
C4$'2005' <- NULL
C4$'2006' <- NULL
C4$'2007' <- NULL
C4$'2008' <- NULL
C4$'2009' <- NULL
C4$'2010' <- NULL
C4$'2011' <- NULL
C4$'2012' <- NULL
C4$'2013' <- NULL

# Create 
C4[, 3] <- round(log(C4[, 3]), digits = 2)
```

```{r echo = FALSE, cache= TRUE, include= TRUE, results= 'asis'}

# Plot perceived opportunties map
Cap1_map <- gvisGeoChart(C4, locationvar = 'iso2c',
                      colorvar = 'mean',
                      options = list(
                        colors = "['#ece7f2', '#2b8cbe']"
))

# Duplicate D4 (perceived capabilities) df
Cap4 <- C4

# Drop iso2c in Dropped4
C4$iso2c <- NULL

# Print the table
print(Cap1_map, tag = 'chart')

```

## Variation in Perceived Opportunities to Start a Business
```{r echo =FALSE, cache=TRUE, include=FALSE, results='asis', warning=FALSE, error=FALSE, message=FALSE}
library("ggplot2")
library("reshape2")

# Create country vector to subset
myvars2 <- c('US','FR','DE','JP','GB')

# Subset Cap4
Cap5 <- subset(Cap4, iso2c %in% myvars2)

# Melt Opp2 df
Cap6 <- melt(data = Cap5,
                  id.vars = c("Country", "iso2c"),
                  measure.vars = c("2001", "2002", "2003", "2004", "2005", "2006",
                                   "2007", "2008", "2009", "2010"))

# Edit column names
names(Cap6) <- gsub("value", "Capabilities", names(Cap6))
names(Cap6) <- gsub("variable", "Year", names(Cap6))

Cap6$iso2c <- NULL
```

```{r echo =FALSE, cache=TRUE, include=TRUE, results='asis', warning=FALSE, error=FALSE, message=FALSE}

# Create plot
ggplot(Cap6,aes(x = Year, y = PerceivedCapabilities, colour = Country, group = Country)) + geom_line()

```

## Variation in Fear Perception in OECD Countries

```{r echo =FALSE, cache=TRUE, include=FALSE, results='asis'}

# Library required packages
library(googleVis)
library(countrycode)
library(repmis)

# Import GEM data frame
PercentFear <- repmis::source_data("https://raw.githubusercontent.com/SamDund/Assign2Education/master/GEMData/GEM_FearFail.csv", header = TRUE, cache = FALSE, na.strings = "-")

## Add country codes and place new column at beginning of the data frame
# Create a vector of country codes
F2 <- countrycode(PercentFear$Country, "country.name", "iso2c")

# Create a new column and insert the country code vector
PercentFear$new.col <- F2

# Reorder columns so country code is first
F3 <- PercentFear[,c(15,1:14)]

# Rename first column
F4 <- plyr::rename(x = F3, 
                   replace = c("new.col" = "iso2c"))

# Remove X's in D4 year columns
names(F4) <- gsub("X", "", names(F4))

# Set columns as numeric
F4[,"2001"] <- as.numeric(F4[,"2001"])
F4[,"2002"] <- as.numeric(F4[,"2002"])
F4[,"2003"] <- as.numeric(F4[,"2003"])
F4[,"2004"] <- as.numeric(F4[,"2004"])
F4[,"2005"] <- as.numeric(F4[,"2005"])
F4[,"2006"] <- as.numeric(F4[,"2006"])
F4[,"2007"] <- as.numeric(F4[,"2007"])
F4[,"2008"] <- as.numeric(F4[,"2008"])
F4[,"2009"] <- as.numeric(F4[,"2009"])
F4[,"2010"] <- as.numeric(F4[,"2010"])
F4[,"2011"] <- as.numeric(F4[,"2011"])
F4[,"2012"] <- as.numeric(F4[,"2012"])
F4[,"2013"] <- as.numeric(F4[,"2013"])

# Save cleaned df
Fear1 <- F4

# Generate a new column with mean of all other columns
F4$mean <- rowMeans(subset(F4, select = c("2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008",
                                          "2009", "2010", "2011", "2012", "2013")), na.rm = TRUE)

# Drop year columns, leaving only mean column
F4$'2001' <- NULL
F4$'2002' <- NULL
F4$'2003' <- NULL
F4$'2004' <- NULL
F4$'2005' <- NULL
F4$'2006' <- NULL
F4$'2007' <- NULL
F4$'2008' <- NULL
F4$'2009' <- NULL
F4$'2010' <- NULL
F4$'2011' <- NULL
F4$'2012' <- NULL
F4$'2013' <- NULL

# Create 
F4[, 3] <- round(log(F4[, 3]), digits = 2)
```

```{r echo = FALSE, cache= TRUE, include= TRUE, results= 'asis'}

# Plot perceived opportunties map
FF_map <- gvisGeoChart(F4, locationvar = 'iso2c',
                      colorvar = 'mean',
                      options = list(
                        colors = "['#ece7f2', '#2b8cbe']"
))

# Duplicate F4 (perceived opportunities) df
DroppedF4 <- F4

# Drop iso2c in Dropped4
DroppedF4$iso2c <- NULL

# Print the table
print(FF_map, tag = 'chart')

```

## Variation in Fear of Failure
```{r echo =FALSE, cache=TRUE, include=FALSE, results='asis', warning=FALSE, error=FALSE, message=FALSE}
library("ggplot2")
library("reshape2")

# Create country vector to subset
myvars2 <- c('US','FR','DE','JP','GB')

# Subset Opp2
Fear2 <- subset(Fear1, iso2c %in% myvars2)

# Melt Opp2 df
Fear3 <- melt(data = Fear2,
                  id.vars = c("Country", "iso2c"),
                  measure.vars = c("2001", "2002", "2003", "2004", "2005", "2006",
                                   "2007", "2008", "2009", "2010"))

# Edit column names
names(Fear3) <- gsub("value", "FearFail", names(Fear3))
names(Fear3) <- gsub("variable", "Year", names(Fear3))

Fear3$iso2c <- NULL
```

```{r echo =FALSE, cache=TRUE, include=TRUE, results='asis', warning=FALSE, error=FALSE, message=FALSE}

# Create plot
ggplot(Fear3,aes(x = Year, y = FearFail, colour = Country, group = Country)) + geom_line()

```

## Varying Perceptions of Good Opportunities to Start a Business in OECD Countries

```{r echo =FALSE, cache=TRUE, include=FALSE, results='asis'}

# Library required packages
library(googleVis)
library(countrycode)
library(repmis)

# Import GEM data frame
PerceivedOpportunities <- repmis::source_data("https://raw.githubusercontent.com/SamDund/Assign2Education/master/GEMData/PerceivedOpportunities.csv", header = TRUE, cache = FALSE, na.strings = "-")

## Add country codes and place new column at beginning of the data frame
# Create a vector of country codes
D2 <- countrycode(PerceivedOpportunities$Country, "country.name", "iso2c")

# Create a new column and insert the country code vector
PerceivedOpportunities$new.col <- D2

# Reorder columns so country code is first
D3 <- PerceivedOpportunities[,c(15,1:14)]

# Rename first column
D4 <- plyr::rename(x = D3, 
                   replace = c("new.col" = "iso2c"))

# Remove X's in D4 year columns
names(D4) <- gsub("X", "", names(D4))

# Set columns as numeric
D4[,"2001"] <- as.numeric(D4[,"2001"])
D4[,"2002"] <- as.numeric(D4[,"2002"])
D4[,"2003"] <- as.numeric(D4[,"2003"])
D4[,"2004"] <- as.numeric(D4[,"2004"])
D4[,"2005"] <- as.numeric(D4[,"2005"])
D4[,"2006"] <- as.numeric(D4[,"2006"])
D4[,"2007"] <- as.numeric(D4[,"2007"])
D4[,"2008"] <- as.numeric(D4[,"2008"])
D4[,"2009"] <- as.numeric(D4[,"2009"])
D4[,"2010"] <- as.numeric(D4[,"2010"])
D4[,"2011"] <- as.numeric(D4[,"2011"])
D4[,"2012"] <- as.numeric(D4[,"2012"])
D4[,"2013"] <- as.numeric(D4[,"2013"])

# Save cleaned df
Opp1 <- D4

# Generate a new column with mean of all other columns
D4$mean <- rowMeans(subset(D4, select = c("2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008",
                                          "2009", "2010", "2011", "2012", "2013")), na.rm = TRUE)

# Drop year columns, leaving only mean column
D4$'2001' <- NULL
D4$'2002' <- NULL
D4$'2003' <- NULL
D4$'2004' <- NULL
D4$'2005' <- NULL
D4$'2006' <- NULL
D4$'2007' <- NULL
D4$'2008' <- NULL
D4$'2009' <- NULL
D4$'2010' <- NULL
D4$'2011' <- NULL
D4$'2012' <- NULL
D4$'2013' <- NULL

# Create 
D4[, 3] <- round(log(D4[, 3]), digits = 2)
```

```{r echo = FALSE, cache= TRUE, include= TRUE, results= 'asis'}

# Plot perceived opportunties map
Opp_map <- gvisGeoChart(D4, locationvar = 'iso2c',
                      colorvar = 'mean',
                      options = list(
                        colors = "['#ece7f2', '#2b8cbe']"
))

# Duplicate D4 (perceived opportunities) df
Dropped4 <- D4

# Drop iso2c in Dropped4
Dropped4$iso2c <- NULL

# Print the table
print(Opp_map, tag = 'chart')

```

## Variation in Perceived Opportunities to Start a Business
```{r echo =FALSE, cache=TRUE, include=FALSE, results='asis', warning=FALSE, error=FALSE, message=FALSE}
library("ggplot2")
library("reshape2")

# Create country vector to subset
myvars2 <- c('US','FR','DE','JP','GB')

# Subset Opp2
Opp2 <- subset(Opp1, iso2c %in% myvars2)

# Melt Opp2 df
Opp3 <- melt(data = Opp2,
                  id.vars = c("Country", "iso2c"),
                  measure.vars = c("2001", "2002", "2003", "2004", "2005", "2006",
                                   "2007", "2008", "2009", "2010"))

# Edit column names
names(Opp3) <- gsub("value", "PerceivedOpportunities", names(Opp3))
names(Opp3) <- gsub("variable", "Year", names(Opp3))

Opp3$iso2c <- NULL
```

```{r echo =FALSE, cache=TRUE, include=TRUE, results='asis', warning=FALSE, error=FALSE, message=FALSE}

# Create plot
ggplot(Opp3,aes(x = Year, y = PerceivedOpportunities, colour = Country, group = Country)) + geom_line()

```

## Regression Results

We decided to ignore country-specific effects because the change in the intercept resulting from fixed effects was small and the slope of the fixed effects and OLS models are very similar.

```{r, echo =FALSE, cache=TRUE, include=FALSE, results='asis', warning=FALSE, error=FALSE, message=FALSE}

library(Zelig)
library(ZeligMultilevel)

# Import CompleteData
CompleteData <- repmis::source_data("https://raw.githubusercontent.com/SamDund/Assign2Education/master/GEMData/CompleteData.csv", header = TRUE)

CompleteData_omit <- na.omit(CompleteData)

# Run a RE regression
Z1 <- zelig(PercentIntention ~ WayofTeaching + PercentPerceived + PercentFear*PercentOpportunities + PercentOpportunities, data = CompleteData_omit, model = "ls")

# Set range of PercentPerceived values in simulation
x.out2 <- setx(Z1, PercentPerceived = 20:70)

# Run simulation (default 1000 trials)
simz2 <- sim(Z1, x = x.out2)

# Adjust margins
par(mar=c(5,5,5,5))

```

```{r, echo =FALSE, cache=FALSE, include=TRUE, results='asis', warning=FALSE, error=FALSE, message=FALSE}
# Plot simulation
plot(simz2, xlab = 'PercentPerceived', ylab = 'PercentIntention | PercentPerceived')
```

## New
```{r, echo =FALSE, cache=TRUE, include=TRUE, results='asis', warning=FALSE, error=FALSE, message=FALSE}

```