### 2014-12-06 ExtraCharts_index/ExtraCharts.Rmd
Removed unnecessary charts and added website text

---
title: "Untitled"
author: "Sam"
date: "December 6, 2014"
output: html_document
---

```{r echo =FALSE, cache=FALSE, include=TRUE, results='asis', warning=FALSE, error=FALSE, message=FALSE}

# Import GEM data frame
PerceivedOpportunities <- repmis::source_data("https://raw.githubusercontent.com/SamDund/Assign2Education/master/GEMData/PerceivedOpportunities.csv", header = TRUE, cache = FALSE, na.strings = "-")

## Add country codes and place new column at beginning of the data frame
# Create a vector of country codes

D2 <- countrycode(PerceivedOpportunities$Country, "country.name", "iso2c")

# Create a new column and insert the country code vector
PerceivedOpportunities$new.col <- D2

# Reorder columns so country code is first
D3 <- PerceivedOpportunities[,c(15,1:14)]

# Rename first column
D4 <- plyr::rename(x = D3, 
                   replace = c("new.col" = "iso2c"))

# Remove X's in D4 year columns
names(D4) <- gsub("X", "", names(D4))

# Melt D4 df
D6 <- melt(data = D4,
                  id.vars = c("Country", "iso2c"),
                  measure.vars = c("2001", "2002", "2003", "2004", "2005", "2006",
                                   "2007", "2008", "2009", "2010"))

# Change column titles to  year and PerceivedOpportunities
names(D6) <- gsub("value", "PerceivedOpportunities", names(D6))    
names(D6) <- gsub("variable", "Year", names(D6))

# Set columns as numeric
D6[,"PerceivedOpportunities"] <- as.numeric(D6[,"PerceivedOpportunities"])

# Save cleaned df
D7 <- D6

# Create country vector to subset
myvars2 <- c('US','FR','DE','JP','GB')

# Subset D6
D8 <- subset(D6, iso2c %in% myvars2)

# Create a scatterplot
D8 %>% 
  ggvis(~Year, ~PerceivedOpportunities) %>%
  layer_points(fill = ~factor(Country)) %>%
  group_by(Country)

D8 %>%
  ggvis(~Year, ~PerceivedOpportunities) %>%
  layer_points(fill = ~factor(Country)) %>%
  group_by(Country)

```