### 2014-12-07 Index_sources/Ameila_source.Rmd
Added Zelig sims

---
title: "Index_sources"
author: "Sam"
date: "December 7, 2014"
output: html_document
---

``````{r echo = FALSE, cache= TRUE, include= FALSE, results= 'asis'}

### Generate a version of "CompleteDataset" with imputed values for additional analysis

CD <- repmis::source_data("https://raw.githubusercontent.com/SamDund/Assign2Education/master/GEMData/CompleteData.csv", header = TRUE, na.strings = "NA")

# Set Year column as a date
CD$Year <- as.Date(as.character(CD$Year), format = "%Y")

# Dropp X column
CD$X <- NULL

# Impute missing values for regression analysis (Amelia)
Cimputed <- Amelia::amelia(CD, m = 5, idvars = c("iso2c", "Country"), ts = "Year")

# Create CSV of imputed values
write.amelia(Cimputed, "Cimputed", separate = TRUE, format = "csv")

```