### 2014-12-12 index.Rmd
Revised charts and removed content included in paper

---
title: 'From Education to Entrepreneurship'
author: "Sam Dundon and Thierry Racaud"
date: "December 6, 2014"
output: html_document
---

<br>

### Key Research Question 
To what extent do teaching methods supporting self-confidence affect the rate of entrepreneurship in OECD countries?

<br>

### Project Overview

We have drawn on publicly available data from the [Global Entrepreneurship Monitor (GEM)](http://www.gemconsortium.org) National Expert Survey (NES) and Adult Population Survey (APS) to create graphic visualizations of participants' responses to surveys about entrepreneurship in [OECD (Organisation for Economic Co-operation and Development)](http://www.oecd.org/) countries.

The following interactive chart can be utilized to see the relationship between the variables we investigated (this can be best viewed in a browser such as [Chrome](https://www.google.com/chrome/)). Among other features, the independent and dependent variables can be selected to see the relationship between each of the variables.  The variables are defined in detail below.

```{r echo =FALSE, cache=FALSE, include=TRUE, results='asis', warning=FALSE, error=FALSE, message=FALSE}
library(ggvis)
library(devtools)
library(googleVis)
library(countrycode)
library(repmis)
library(reshape2)
library(shiny)
library(httpuv)

# Create a vector of packages used
pkgs <- c('repmis', 'ggvis', 'devtools', 'googleVis', 'countrycode', 'reshape2', 'shiny', 'httpuv', 'Zelig',
          'stargazer', 'Amelia')

# Create citations txt file
repmis::LoadandCite(pkgs, file= 'RpackageCitations.bib')

# Import "CompleteData" df from repository
CD <- repmis::source_data("https://raw.githubusercontent.com/SamDund/Assign2Education/master/GEMData/CompleteData.csv", header = TRUE, na.strings = "NA")

names(CD) <- gsub("PercentPerceived", "PerceivedCapability", names(CD))

# Set Year column as a date
CD$Year <- as.Date(as.character(CD$Year), format = "%Y")

CD_copy <- CD

# Drop X column
CD$X <- NULL
CD$iso2c <- NULL

R <- gvisMotionChart(CD, idvar = "Country", timevar = "Year", options = list(state='{"showTrails":false};'))

print(R, tag = 'chart')
```

<br>

### Variables:

1. **Entrepreneurial Intention**: Percentage of 18-64 population (individuals involved in any stage of entrepreneurial activity excluded) who intend to start a business within three years.

2. **Nascent Entrepreneurship**: Percentage of 18-64 population who are currently a nascent entrepreneur, i.e., actively involved in setting up a business they will own or co-own; this business has not paid salaries, wages, or any other payments to the owners for more than three months.

3. **Established Entrepreneur**: Percentage of 18-64 population who are currently owner-manager of an established business, i.e., owning and managing a running business that has paid salaries, wages, or any other payments to the owners for more than 42 months.

4. **Way of Teaching**: In my country, teaching in primary and secondary education encourages creativity, self-sufficiency, and personal initiative.

5. **Perceived Capabilities**: Percentage of 18-64 population who believe to have the required skills and knowledge to start a business

6. **Fear of Failure**: Percentage of 18-64 population with positive perceived opportunities who indicate that fear of failure would prevent them from setting up a business.

7. **Perceived Opportunities**: Percentage of 18-64 population who see good opportunities to start a firm in the area where they live

<br>

### Causal Chain

Our decision to select and investigate these variables was based on the following causal chain

1. An education system that reinforces self-confidence

2. Higher degree of individual self-esteem

3. Higher probability an individual becomes an entrepreneur

4. Higher rate of entrepreneurship at a country-level

<br>

### Geographic Variation in our Explanatory Variable

As expected, there is significant variation among OECD countries with respect to each of the variables investigated.  The most important of these variables is our explanatory variable, "WayofTeaching".  This is a measure of how well teaching methods support self-confidence in each country.  The following interactive chart displays variation in self confidence with the log of the mean survey response from each country over 2001-2013 time frame.

```{r echo =FALSE, cache=FALSE, include=TRUE, results='asis', warning=FALSE, error=FALSE, message=FALSE}
# Import GEM data frame
WayofTeaching <- repmis::source_data("https://raw.githubusercontent.com/SamDund/Assign2Education/master/GEMData/WayofTeaching", header = TRUE, cache = FALSE, na.strings = "NA")

# Drop unnecessary column 1
WayofTeaching$X <- NULL

# Change data from long to wide format
WT <- reshape(WayofTeaching, idvar = c("iso2c", "Country"), timevar = "Year", direction = "wide")

# Delete variable name in Year columns
names(WT) <- gsub("WayofTeaching.", "", names(WT))

# Duplicate df for future use
WT2 <- WT

# Generate a new column with mean of all other columns
WT$mean <- rowMeans(subset(WT, select = c("2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008",
                                          "2009", "2010")), na.rm = TRUE)

# Drop year columns, leaving only mean column
WT$'2001' <- NULL
WT$'2002' <- NULL
WT$'2003' <- NULL
WT$'2004' <- NULL
WT$'2005' <- NULL
WT$'2006' <- NULL
WT$'2007' <- NULL
WT$'2008' <- NULL
WT$'2009' <- NULL
WT$'2010' <- NULL
WT$'2011' <- NULL
WT$'2012' <- NULL
WT$'2013' <- NULL
WT$row.names <- NULL

# Create 
WT[, 3] <- round(log(WT[, 3]), digits = 2)
```

```{r echo = FALSE, cache= TRUE, include= TRUE, results= 'asis'}
library(googleVis)

# Plot
WT_map <- gvisGeoChart(WT, locationvar = 'iso2c',
                      colorvar = 'mean',
                      options = list(
                        colors = "['#ece7f2', '#2b8cbe']"
))

print(WT_map, tag = 'chart')

```

<br>

### Statistical Analysis

The variables indicated above were used to obtain inferential statistics:

Variable | Indicator 
-------------|---------------
Y Dependent Variable | Rate of Entrepreneurship (Intention, Nascent, Established)
X1 Explanatory Variable | Way of Teaching
X2 Control Variable | Perceived Capabilities
X3 Control Variable | Fear of Failure / Perceived Opportunities
X4 Control Variable | Perceived Opportunities

We have included our findings in a detailed paper available here.  In sum, however, the most statistically significant and important findings from our analysis are that there is some evidence of a postitive relationship between higher levels of self-confidence reinforcement in teaching (a higher "WayofTeaching" score) and the rate of people owning their own business.  This can be seen in the chart below.  As we can see, however, when we add a trendline for each country the relationship becomes less certain.  For example, in the case of the United States and the United Kingdom, the country-specific trend line has a decreasing slope.  We believe this may be due to inconsistency in the survey and that a larger data set would allow us to more accurately describe the relationship between these variables.

<br>

```{r echo = FALSE, cache= FALSE, include= TRUE, results='asis', warning=FALSE}

# Create country vector to subset
myvars2 <- c('US','FR','DE','JP','GB')

# Subset CD
CD_subset <- subset(CD_copy, iso2c %in% myvars2)

CD_subset2 <- subset(CD_copy, iso2c %in% myvars2)

# Remove unecessary columns
CD_subset$X <- NULL
CD_subset$iso2c <- NULL
CD_subset$PercentFear <- NULL
CD_subset$PercentNascent <- NULL
CD_subset$PercentPerceived <- NULL
CD_subset$PercentIntention <- NULL
CD_subset$PercentOpportunities <- NULL

# Create plot of WoT and PercentOwner
CD_subset %>% 
  ggvis(x = ~WayofTeaching, y = ~PercentOwner) %>% 
  layer_points(fill = ~factor(Country)) %>%
  group_by(Country)
```

<br>

In addiiton, we can see a positive relationship between the percentage of respondents who believe there are opportunities to establish their own business (the percentage of people perceiving opportunities) and the rate of business ownership as reported in the survey.  This relationship can be seen in the following chart.  Similar to our findings above, this chart seems to show an overall positive relationship, but looking at country-specific factors, we can see that data from Japan indicates a negative slope.  We believe access to additional data would allow us to more accurately describe the relationship between these variables and then to investigate why these differences exist.

<br>

```{r echo = FALSE, cache= FALSE, include= TRUE, results='asis', warning=FALSE}

# Drop unnecesssary columns in copied df
CD_subset2$X <- NULL
CD_subset2$iso2c <- NULL
CD_subset2$PercentFear <- NULL
CD_subset2$PercentNascent <- NULL
CD_subset2$PercentOpportunity <- NULL
CD_subset2$PercentIntention <- NULL
CD_subset2$WayofTeaching <- NULL

names(CD_subset2) <- gsub("PercentPerceived", "PerceivedCapability", names(CD_subset2))

# Create plot of PercentOpp and PercentOwner
CD_subset2 %>% 
  ggvis(x = ~PerceivedCapability, y = ~PercentOwner) %>% 
  layer_points(fill = ~factor(Country)) %>%
  group_by(Country)

```

### Conclusions

Significant work remains to more clearly understand the extent of the relationship between teaching methods supporting self-confidence and a country's rate of entrepreneurship.  This research, however, does show that a positive relationship exists.  One clear requirement of further analysis in this field is increased availability and consistency in collected data.  Missing data generates substantial uncertainty and limits our ability to strongly state conclusions.  This will need to be addressed for future projects.