### 2014-12-12 MiscData/Misc/MiscSims.Rmd
Re-arranged extra repository data

---
title: "Trial Simulations"
author: "Sam"
date: "November 14, 2014"
output: html_document
---

The second simulation was

```{r echo = FALSE, cache=TRUE, include=TRUE, results='asis' }

# Set range of PercentFear values in simulation
x.out <- setx(Z1, PercentFear = 20:100)

# Run simulation (default 1000 trials)
simz1 <- sim(Z1, x = x.out)

# Adjust margins
par(mar=c(2,3,2,2))

# Plot simulation
plot(simz1)

```

```{r echo = FALSE, cache=TRUE, include=TRUE, results='asis' }

# Set range of PercentOpportunities values in simulation
x.out3 <- setx(Z1, PercentOpportunities = 20:100)

# Run simulation (default 1000 trials)
simz3 <- sim(Z1, x = x.out3)

# Adjust margins
par(mar=c(2,3,2,2))

# Plot simulation
plot(simz3)

```